---
title: "Group1_RCode"
output: pdf_document
date: '2022-06-25'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Group 1 R Code

Brandon Coates and Jessica Bunge

**Load Libraries**

```{r}
#Load Libraries
library(TSA)
library(ggplot2)
library(ggpubr)
library(zoo)
library(xts)
library(tsbox)
library(tseries)
library(forecast)
library(sandwich)
library(lmtest)
library(AICcmodavg)
```

## Data Loading, Cleaning, and Exploration

Import the data into R:

```{r}
#Load data
climatetrain <- read.csv("DailyDelhiClimateTrain.csv",
                      stringsAsFactors = FALSE)
climatetest<- read.csv("DailyDelhiClimateTest.csv",
                       stringsAsFactors = FALSE)
```

Next perform data cleaning:

```{r}
#Remove last row of climate train data 
climate <- climatetrain[-c(1462),]

#Remove unnecessary variables
climate <- climate[,-c(3, 4, 5)]
climatetest <- climatetest[,-c(3, 4, 5)]

#Convert dates to date-time
#Use climatetest$date <- as.Date(climatetest$date, "%m/%d/%Y")
#If error occurs during xts reformatting
climate$date <- as.Date(climate$date)
class(climate$date)

climatetest$date <- as.Date(climatetest$date)
class(climatetest$date)
```

Create initial time series plot:

```{r pressure, echo=FALSE}
#Plot time series plot
temp <- qplot(x = date, y = meantemp, data = climate,
      xlab = "Date", ylab = "Mean Temperature (Celsius)",
      main = "Mean Daily Temperatures in Delhi")
temp
```

Convert data into time series format (xts):

```{r}
#Convert to xts data format
climateXTS <- xts(climate[,-1], order.by = climate$date)
climatetestXTS <- xts(climatetest[,-1], order.by = climatetest$date)
```

Convert daily temperature records into average monthly temperatures:

```{r}
#Find mean temperature for each month
climatemon <- apply.monthly(climateXTS, mean)
climatetestmon <- apply.monthly(climatetestXTS, mean)
head(climatemon)
```

Now convert from xts object format to data frame format:

```{r}
#Transform to data frame
climatemonth <- ts_df(climatemon)
climatetestmonth <- ts_df(climatetestmon)
head(climatemonth)
```

Now split the data into training and test set (42 months/9 months) and convert to time series object:

```{r}
#Assign training and testing set data
climatetestmonth <- rbind(climatemonth[43:48,], climatetestmonth)
climatemonth <- climatemonth[-(43:48),]

#Convert data frames to time series objects
climatemonth <- ts(as.vector(climatemonth$value), start=c(2013,1), end = c(2016, 6), frequency = 12)
climatetest <- ts(as.vector(climatetestmonth$value), start=c(2016,7), end = c(2017, 3), frequency = 12)
```

Use the decompose function on the training set to observe the trend components of the model:

```{r}
#Use decompose to examine components of data series
decomp1 <- decompose(climatemonth)
plot(decomp1)
```

### Check Transformations

To check if log transformation is necessary:

```{r}
#Check if log transformation is appropriate
par(mfrow = c(2,1))
plot(climatemonth, main = "Original")
abline(lm(climatemonth~time(climatemonth)))
plot(log(climatemonth), main = "Log Transformed")
abline(lm(log(climatemonth)~time(log(climatemonth))))
```

To check if differencing is necessary:

```{r}
#Differenced vs log-differenced
plot(diff(climatemonth), main = "Differenced")
plot(diff(log(climatemonth)), main = "Log difference")
```

## Model Building

### Seasonal Means Model with Linear Compnent

To build seasonal means model with a linear component:

```{r}
#Try seasonal means model
x <- climatemonth
tm <- time(x)
month <- season(x)
model1 <- lm(x ~ tm + month)
summary(model1)
```

To check the residuals plot of the model:

```{r}
#Residual plot for linear component and seasonal means model
par(mfrow = c(1,1))
lx1 <- ts(resid(model1), start = c(2013, 1), frequency = 12)
plot(lx1, main = "Residuals of Seasonal Means and Linear Trend Model", type = 'o')
abline(h=0, lty = 2)
```

Plot the time series data with the model fit line:

```{r}
# data vs fitted line
plot(x)
f.lx1=ts(fitted(model1),start=c(2013,1),freq=12)
lines(f.lx1,col=2,lty=2)
```

Perform HAC test on components:

```{r}
#Check significance of variables
coeftest(model1, vcov=vcovHAC(model1))
```

#### Model Diagnostics

To check the residuals:

```{r}
# 1) Residual plot for seasonal means model (zero mean and homoscedasticity)
plot(rstandard(model1),type="l")
abline(h=0,col=2)
```

To check normality via QQ plot:

```{r}
# 2) QQ plot for seasonal means model (normality)
qqnorm(lx1)
qqline(lx1)
```

Shapiro-Wilk Test and Runs Test:

```{r}
# 3) Shapiro-Wilk test (normality) and runs test (independence) for seasonal means model
shapiro.test(lx1)
runs(lx1)
```

To check the ACF plot:

```{r}
# 4) ACF plot for seasonal means model(independence)
Acf(lx1)
```

### Seasonal Means with Non-Linear Trend Component

The seasonal means model with a polynomial is created and plotted with the data set:

```{r}
#Try nonlinear trend with seasonal means model
tm=time(x)
tm2=time(x)^2
model2=lm(x~month+tm+tm2)
plot(x)
f.lx2=ts(fitted(model2),start=c(2013,1),freq=12)
lines(f.lx2,col=2,lty=2)
lx2=residuals(model2)
```

#### Model Diagnostics

The residual plot for this model is:

```{r}
# 1) Residual plot (zero mean and homoscedasticity)
plot(rstandard(model2),type="o", main = "Non-Linear Trend")
abline(h=0,col=2)
```

The QQ plot for the model to check normality:

```{r}
# 2) QQ plot (normality)
qqnorm(lx2)
qqline(lx2)
```

The Shapiro-Wilk Test and Runs Test results:

```{r}
# 3) Shapiro-Wilk test (normality) and runs test (independence)
shapiro.test(lx2)
runs(lx2)
```

The ACF plot of the residuals

```{r}
# 4) ACF plot (independence)
Acf(lx2)
#Nonlinear trend does not appear to be necessary for this model
```

### ARIMA Model Building

Using residuals from seasonal means with linear component model, the need for differencing was examined:

```{r}
#d = 0
plot(lx1,type="l")
Acf(lx1)
adf.test(lx1)
pp.test(lx1)
```

Order for p and q was examined next:

```{r}
# p & q
par(mfrow=c(1,2))
Acf(lx1)
Pacf(lx1)
par(mfrow=c(1,1))
eacf(lx1)

#p=1 and q=1 were found to be best
```

This suggested MA(1), AR(1), and ARMA(1,1)

#### MA(1) model

The MA(1) model was built:

```{r}
ma1 <- Arima(lx1, order = c(0,0,1), include.mean = FALSE)
```

The summary of the MA(1) model:

```{r}
ma1
```

#### AR(1) Model

The AR(1) model was built:

```{r}
ar1 <- Arima(lx1, order = c(1,0,0), include.mean = FALSE)
```

The summary of the model:

```{r}
ar1
```

#### ARMA(1,1)

The ARMA(1,1) model was built:

```{r}
arma11 <- Arima(lx1, order = c(1,0,1), include.mean = FALSE)
```

The summary of the model:

```{r}
arma11
```

### SARIMA Model

The auto arima function was used to identify a candidate SARIMA model:

```{r}
#auto arima for SARIMA
mo <- lm(x ~ tm)
f0 <- ts(resid(mo), start = c(2013, 1), frequency = 12)

#Fit Auto Arima Model
aarima <- auto.arima(f0)

```

The summary of the SARIMA(1,0,0)(1,1,0)[12] model was:

```{r}
aarima
```

#### Model Diagnostics

The residual plot of the SARIMA model is:

```{r}
#Residual plot for SARIMA
plot(rstandard(aarima), type = 'o', main = "Residuals of SARIMA Model")
abline(h = 0, col = 2)
```

The QQ plot to assess normality:

```{r}
qqnorm(rstandard(aarima), main = "Q-Q Plot of Residuals")
qqline(rstandard(aarima))
```

Shapiro-Wilk Test and Runs Test for SARIMA model:

```{r}
shapiro.test(rstandard(aarima))
runs(rstandard(aarima))
```

### Candidate Models

Two candidate models were chosen, the AR(1) model from the seasonal means model with the linear component and the SARIMA model

#### AR(1) Model Diagnostics

The diagnostics plot for the AR(1) candidate model:

```{r}
tsdiag(ar1)
```

#### SARIMA(1,0,0)(1,1,0)[12] Model Diagnostics

The diagnostics plots for the SARIMA model are:

```{r}
tsdiag(aarima)
```

### Model Downselection

Over and underfitting was conducted to find the optimum model for the data set. For the AR(1) model:

```{r}
#Over and under fitting
#AR1 model
ar2<- Arima(lx1, order = c(2, 0, 0), include.mean = FALSE)
ar2 #additional parameter is not significant
```

For the SARIMA(1,0,0)(1,1,0)[12] model:

```{r}
#SARIMA (1,0,0)(1,0,0)12 model
mod <- Arima(f0,order=c(1,0,0),seasonal=list(order=c(1,0,0)), include.mean = FALSE)
mod
mod1 <- Arima(f0,order=c(1,0,0),seasonal=list(order=c(1,1,0)), include.mean = FALSE)
mod1
mod2 <- Arima(f0,order=c(1,0,0),seasonal=list(order=c(1,1,1)), include.mean = FALSE)
mod2 #Found to be best choice
mod3 <- Arima(f0, order=c(1,1,0),seasonal=list(order=c(1,0,0)), include.mean = FALSE)
mod3
mod4 <- Arima(f0,order=c(1,1,0),seasonal=list(order=c(1,1,0)), include.mean = FALSE)
mod4
mod5 <- Arima(f0,order=c(1,1,0),seasonal=list(order=c(1,1,1)), include.mean = FALSE)
mod5
mod6 <- Arima(f0,order=c(1,1,1),seasonal=list(order=c(1,0,0)), include.mean = FALSE)
mod6
mod7 <- Arima(f0,order=c(1,1,1),seasonal=list(order=c(1,1,0)), include.mean = FALSE)
mod7
mod8 <- Arima(f0,order=c(1,1,1),seasonal=list(order=c(1,1,1)), include.mean = FALSE)
mod8
mod9 <- Arima(f0,order=c(2,0,0),seasonal=list(order=c(1,1,1)), include.mean = FALSE)
mod9
mod10 <- Arima(f0,order=c(2,1,0),seasonal=list(order=c(1,1,1)), include.mean = FALSE)
mod10
mod11 <- Arima(f0,order=c(2,0,1),seasonal=list(order=c(1,1,1)), include.mean = FALSE)
mod11
mod12 <- Arima(f0,order=c(1,0,0),seasonal=list(order=c(2,1,1)), include.mean = FALSE)
mod12
```

The final model was chosen to be SARIMA(1,0,0)(1,1,1)[12] model

### Final Model Summary and Diagnostics

The SARIMA(1,0,0)(1,1,1)[12] model was assigned to be the final model:

```{r}
#Assign chosen final model and check diagnostics
modfinal <- mod2
modfinal
```

The diagnostics plots for this model were:

```{r}
tsdiag(modfinal)
```

## Forecasting

### Forecasting on Test Set

The final model was used to forecast the values of the test set:

```{r}
#Forecasting on test set
newtm <- seq(from=2016.583, to=2017.25, length = 9)
newmonth <- as.factor(c("July", "August", "September", "October",
                            "November", "December", "January", "February", "March"))
newdata <- data.frame(month=newmonth, tm=newtm)
predxreg <- predict(mo, newdata)

predx <- predict(modfinal, n.ahead = 9)
pr <- predx$pred + predxreg
uci <- pr + 2 * predx$se
lci <- pr - 2 * predx$se

#Plot forecast
pr <- ts(pr, start = c(2016,7), end = c(2017, 3), frequency = 12)
uci <- ts(uci, start = c(2016,7), end = c(2017,3), frequency = 12)
lci <- ts(lci, start = c(2016,7), end = c(2017,3), frequency = 12)

ymin <- min(c(as.vector(lci), x))-.1
ymax <- max(c(as.vector(uci), x))+.1

par(mfrow=c(1,1))
plot(x, xlim = c(2013, 2017.5), ylim = c(0, 40), main = "Mean Monthly Temperature of New Delhi",
     ylab = "Mean Temperature (Degrees Celsius)")
lines(pr, col=5)
lines(uci, col=3, lty = 2)
lines(lci, col=3, lty =2)
lines(climatetest, col = 2)
abline(h = 10, lty = 2)
legend(2016, 12, legend = c("Forecast", "Test Set", "Upper 95%", "Lower 95%"),
       col=c(5, 2, 3, 3), lty=c(1, 1, 2, 2), cex=0.8)
       
```

The forecast accuracy was then calculated:

```{r}
#Calculate forecast Accuracy
accuracy(pr, climatetest)
```

### Forecasting Mean Temperatures for 10 Years

The model was then used to forecast mean monthly temperatures until 2027:

```{r}
#Forecast mean temperatures 10 years out
newtm10 <- seq(from=2016.583, to=2027, length = 126)
testmonth <- as.factor(c("July", "August", "September", "October",
                        "November", "December"))
newmonth <- as.factor(rep(c("January", "February", "March", "April", "May",
                            "June", "July", "August", "September", "October",
                            "November", "December"), times = 10))
month10 <- append(testmonth,newmonth)
newdata10 <- data.frame(month=month10, tm=newtm10)
predxreg10 <- predict(mo, newdata10)

predx10 <- predict(modfinal, n.ahead = 126)
pr10 <- predx10$pred + predxreg10
uci10 <- pr10 + 2 * predx10$se
lci10 <- pr10 - 2 * predx10$se

#Plot 10 year forecast
pr10 <- ts(pr10, start = c(2016,7), end = c(2027, 1), frequency = 12)
uci10 <- ts(uci10, start = c(2016,7), end = c(2027,1), frequency = 12)
lci10 <- ts(lci10, start = c(2016,7), end = c(2027,1), frequency = 12)

ymin <- min(c(as.vector(lci10), x))-.1
ymax <- max(c(as.vector(uci10), x))+.1

plot(x, xlim = c(2013, 2027.5), ylim = c(10,50), main = "Mean Monthly Temperature of New Delhi",
     ylab = "Mean Temperature (Degrees Celsius)")
lines(pr10, col=2)
lines(uci10, col=3, lty = 2)
lines(lci10, col=3, lty = 2)
abline(h = 10, lty = 2)
legend(2023, 18, legend = c("Forecast", "Upper 95%", "Lower 95%"),
       col=c(2, 3, 3), lty=c(1, 2, 2), cex=0.8)
```

###  Forecasting Mean Temperatures for 20 Years

The model was used to forecast mean monthly temperatures until 2037:

```{r}
#Forecast mean temperatures for next 20 years
newtm20 <- seq(from=2016.583, to=2037, length = 246)
testmonth <- as.factor(c("July", "August", "September", "October",
                         "November", "December"))
newmonth <- as.factor(rep(c("January", "February", "March", "April", "May",
                            "June", "July", "August", "September", "October",
                            "November", "December"), times = 20))
month20 <- append(testmonth,newmonth)
newdata20 <- data.frame(month=month20, tm=newtm20)
predxreg20 <- predict(mo, newdata20)

predx20 <- predict(modfinal, n.ahead = 246)
pr20 <- predx20$pred + predxreg20
uci20 <- pr20 + 2 * predx20$se
lci20 <- pr20 - 2 * predx20$se

#Plot 20 year forecast
pr20 <- ts(pr20, start = c(2016,7), end = c(2037, 1), frequency = 12)
uci20 <- ts(uci20, start = c(2016,7), end = c(2037,1), frequency = 12)
lci20 <- ts(lci20, start = c(2016,7), end = c(2037,1), frequency = 12)

ymin <- min(c(as.vector(lci20), x))-.1
ymax <- max(c(as.vector(uci20), x))+.1

plot(x, xlim = c(2013, 2037.5), ylim = c(10,60), main = "Mean Monthly Temperature of New Delhi",
     ylab = "Mean Temperature (Degrees Celsius)")
lines(pr20, col=2)
lines(uci20, col=3, lty = 2)
lines(lci20, col=3, lty = 2)
abline(h = 10, lty = 2)
legend(2030, 20, legend = c("Forecast", "Upper 95%", "Lower 95%"),
       col=c(2, 3, 3), lty=c(1, 2, 2), cex=0.8)

```
